Realizability-Preserving Two-Moment DG Method for Fermionic Transport \tnoteref{ldrd-support}

### 2018-01-26 Documents/M1/realizableFermionicM1/sections/realizability.tex
Finished the proof of Lemma 1.

\begin{define}
  The moments $\vect{\cM}=\big(\cJ,\vect{\cH}\big)^{T}$ are realizable if they are consistent with a distribution function satisfying $0\le f \le 1$.
\end{define}
\begin{lemma}
  Let the distribution function $f$ be bounded by $0\le f \le1$.  
  Then $0\le\cJ\le1$ and $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$
\end{lemma}
\begin{proof}
  Since this lemma is independent from $\vect{z}$ and t, we can drop them from the expression for convenience.
  Let's proof $0\le\cJ\le1$.

@@

    For $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$, we can proof it's valid with 1-dimension $\vect{\cH}$ first.
    In 1-dimension, $\vect{\ell}(\omegaNu) = \omegaNu$ and
    \begin{equation}
    \vect{\cH}
      =\f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu) \omegaNu d\omegaNu = \f{1}{2}\int_{-1}^{1}f(\omegaNu) \omegaNu d\omegaNu.
    \end{equation}
    Following Banach \& Larecki \cite{banachLarecki_2017}, we can rewrite it to be
    \begin{align}
    \vect{\cH}
        &=(1-\cJ)\,\cJ \nonumber \\
        &-\Big\{\,
          \f{1}{2}\int_{-1}^{1-2\cJ}(1-2\cJ-\omegaNu)f(\omegaNu)\,d\omegaNu
          +\f{1}{2}\int_{1-2\cJ}^{1}(2\cJ-1+\omegaNu)[1-f(\omegaNu)]\,d\omegaNu
        \,\Big\}.
        \label{eq:momentumRewrite1}
    \end{align}
    Since $(1-2\cJ-\omegaNu)f(\omegaNu)$ is positive when $-1\leq\omegaNu\leq1-2\cJ$ and $(2\cJ-1+\omegaNu)[1-f(\omegaNu)]$ is positive when $1-2\cJ\leq\omegaNu\leq1$,
    \begin{align}
    \vect{\cH} \leq (1-\cJ)\,\cJ.
    \label{eq:HupBry1D}
    \end{align}
    Similarly, we can also rewrite $\vect{\cH}$ as
    \begin{align}
     \vect{\cH}
        &=-(1-\cJ)\,\cJ \nonumber \\
        &-\Big\{\,
          \f{1}{2}\int_{-1}^{2\cJ-1}(1-2\cJ+\omegaNu)[1-f(\omegaNu)]\,d\omegaNu
          +\f{1}{2}\int_{2\cJ-1}^{1}(2\cJ-1-\omegaNu)f(\omegaNu)\,d\omegaNu
        \,\Big\}.
        \label{eq:momentumRewrite2}
    \end{align}
    Since $(1-2\cJ+\omegaNu)[1-f(\omegaNu)]$ is negative when $-1\leq\omegaNu\leq2\cJ-1$ and $(2\cJ-1-\omegaNu)f(\omegaNu)$ is negative when $2\cJ-1\leq\omegaNu\leq1$,
    \begin{align}
    \vect{\cH} \geq -(1-\cJ)\,\cJ.
     \label{eq:HlowBry1D}
    \end{align}
    Combining Eq.~\eqref{eq:HupBry1D} and ~\eqref{eq:HlowBry1D} gives $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$ in 1-dimension.
\end{proof}

>>>>

\begin{define}
  The moments $\vect{\cM}=\big(\cJ,\vect{\cH}\big)^{T}$ are realizable if they are consistent with a distribution function satisfying $0\le f \le 1$.
\end{define}
\begin{lemma}
  Let the distribution function $f$ be bounded by $0\le f \le1$.  
  Then $0\le\cJ\le1$ and $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$
\end{lemma}
\begin{proof}
  Since this lemma is independent from $\vect{z}$ and t, we can drop $\vect{z}$ and t from the expressions for this proof.
  Let's prove $0\le\cJ\le1$.

@@

    For $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$, we can prove it's valid with 1-dimension $\vect{\cH}$ first.
    In 1-dimension, $\vect{\ell}(\omegaNu) = \omegaNu$ and
    \begin{equation}
    \vect{\cH}
      =\f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu) \omegaNu d\omegaNu = \f{1}{2}\int_{-1}^{1}f(\omegaNu) \omegaNu d\omegaNu.
    \end{equation}
    Following Banach \& Larecki \cite{banachLarecki_2017}, we can rewrite it to be
    \begin{align}
    \vect{\cH}
        &=(1-\cJ)\,\cJ \nonumber \\
        &-\Big\{\,
          \f{1}{2}\int_{-1}^{1-2\cJ}(1-2\cJ-\omegaNu)f(\omegaNu)\,d\omegaNu
          +\f{1}{2}\int_{1-2\cJ}^{1}(2\cJ-1+\omegaNu)[1-f(\omegaNu)]\,d\omegaNu
        \,\Big\}.
        \label{eq:momentumRewrite1}
    \end{align}
    Since $(1-2\cJ-\omegaNu)f(\omegaNu)$ is positive when $-1\leq\omegaNu\leq1-2\cJ$ and $(2\cJ-1+\omegaNu)[1-f(\omegaNu)]$ is positive when $1-2\cJ\leq\omegaNu\leq1$,
    \begin{align}
    \vect{\cH} \leq (1-\cJ)\,\cJ.
    \label{eq:HupBry1D}
    \end{align}
    Similarly, we can also rewrite $\vect{\cH}$ as
    \begin{align}
     \vect{\cH}
        &=-(1-\cJ)\,\cJ \nonumber \\
        &-\Big\{\,
          \f{1}{2}\int_{-1}^{2\cJ-1}(1-2\cJ+\omegaNu)[1-f(\omegaNu)]\,d\omegaNu
          +\f{1}{2}\int_{2\cJ-1}^{1}(2\cJ-1-\omegaNu)f(\omegaNu)\,d\omegaNu
        \,\Big\}.
        \label{eq:momentumRewrite2}
    \end{align}
    Since $(1-2\cJ+\omegaNu)[1-f(\omegaNu)]$ is negative when $-1\leq\omegaNu\leq2\cJ-1$ and $(2\cJ-1-\omegaNu)f(\omegaNu)$ is negative when $2\cJ-1\leq\omegaNu\leq1$,
    \begin{align}
    \vect{\cH} \geq -(1-\cJ)\,\cJ.
     \label{eq:HlowBry1D}
    \end{align}
    Combining Eq.~\eqref{eq:HupBry1D} and ~\eqref{eq:HlowBry1D} gives $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$ in 1-dimension.

    To prove $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$ for 3-dimensional $\vect{\cH}$, let's introduce a unit vector $\bn$ with a arbitrary direction.
    Then 
    \begin{equation}
    \vect{\cH} \cdot \bn = |\vect{\cH}| \cos\alpha,
    \end{equation}
    with $\alpha$ be the angle between $\vect{\cH}$ and $\bn$. 
    We know that $-|\vect{\cH}|\leq\vect{\cH} \cdot \bn\leq|\vect{\cH}|$ since $-1\leq\cos\alpha\leq1$.
    While given by definition,
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) d\omegaNu \cdot \bn \nonumber \\
     &= \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) \cdot \bn d\omegaNu \nonumber \\
     &= \f{1}{4\pi}\int_{0}^{\pi}\int_{0}^{2\pi}f(\theta,\phi)\ \cos \theta \sin\theta d\theta d\phi
    \end{align}
    while $\theta$ is the angle between $\bl$ and $\bn$, and the coordinate are rotated so that the $\theta$ from $\omegaNu(\theta,\phi)$ is the same with the angle between $\bl$ and $\bn$.
    Note the above expression can also be rewritten by changing variables as
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{2}\int_{0}^{\pi} \f{1}{2\pi}\int_{0}^{2\pi}f(\theta,\phi)\ d\phi \cdot \cos \theta \sin\theta d\theta  \\
    & =  \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi \mu d\mu.
    \end{align}
    Besides,
    \begin{align}
    \cJ = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\,d\omegaNu = \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi d\mu.
    \end{align}
    If we define $f(\mu) =  \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi$, then
    \begin{align}
    \cJ &= \f{1}{2}\int_{-1}^{1} f(\mu) d\mu, \\
    \vect{\cH} \cdot \bn  &= \f{1}{2}\int_{-1}^{1} f(\mu) \mu d\mu.
    \end{align}
    We know $0\leq\,f(\mu)\leq1$.
    Recall the proof we gave for 1-dimension $\vect{\cH}$, we have
    \begin{equation}
     \big(1-\cJ\big)\cJ-\vect{\cH} \cdot \bn\ge0.
    \end{equation}
    Note this inequality is valid for all the possible $\bn$ which includes the special case with $\cos\alpha=1$ and
    \begin{equation}
    \big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0.
    \end{equation} 
\end{proof}
